# census.Rmd

Create the output file census.csv

Files read.

```{r}
path.census.in <- "../data/raw/neighborhood-data/census.csv"

```

Files written.

```{r}
dir.output="../data/v6/output/"
path.census.out <- paste(dir.output,
                         "census.csv",
                         sep="")
```

## Control the script

Set script control variables.

```{r}
testing <- FALSE
```

Initialize R.

```{r}
options(warn=2)  # turn warnings into errors
set.seed(1)      # random number generator seed
require("compiler")
enableJIT(3)     # 3 ==> maximum JIT level
```

Source files here, now that the JIT level is set

```{r}
# nothing to source
```

## Read data

```{r}
# skip first line, as it contains hard-to-interpret feature names
# and easy-to-interpret feature names are in the second line
df <- read.table(path.census.in,
                 header=TRUE,
                 skip=1, 
                 sep="\t",
                 comment.char="",
                 stringsAsFactors=FALSE,
                 na.strings="")
head(df)
```

## Isolate the census tract feature

```{r}
geo.id2 <- df$Geography.Identifier2
fips.code <- floor(geo.id2/1000000)
census.tract <-geo.id2 - fips.code * 1000000

floor(geo.id2)[1:10]
census.tract[1:10]
```

## Create new feature average commute time

```{r}
not.home = "Workers.16.years.and.over..Did.not.work.at.home"
workers.not.at.home <- df[[not.home]]

time <- function(specification) {
  result <- paste(not.home,
                  "..Travel.time.to.work..",
                  specification,
                  ".minutes",
                  sep="")
  result
  }

commute.less.5 <- df[[time("Less.than.5")]]
commute.5.to.9 <- df[[time("5.to.9")]]
commute.10.to.14 <- df[[time("10.to.14")]]
commute.15.to.19 <- df[[time("15.to.19")]]
commute.20.to.24 <- df[[time("20.to.24")]]
commute.25.to.29 <- df[[time("25.to.29")]]
commute.30.to.34 <- df[[time("30.to.34")]]
commute.35.to.39 <- df[[time("35.to.39")]]
commute.40.to.44 <- df[[time("40.to.44")]]
commute.45.to.59 <- df[[time("45.to.59")]]
commute.60.to.89 <- df[[time("60.to.89")]]
commute.more.90 <- df[[time("90.or.more")]]

avg.commute <-
  ((commute.less.5 * ((0 + 5) / 2)) +
     (commute.5.to.9 * ((5 + 9) / 2)) +
     (commute.10.to.14 * ((10 + 14) / 2)) +
     (commute.15.to.19 * ((15 + 19) / 2)) +
     (commute.20.to.24 * ((20 + 24) / 2)) +
     (commute.25.to.29 * ((25 + 29) / 2)) +
     (commute.30.to.34 * ((30 + 34) / 2)) +
     (commute.35.to.39 * ((35 + 39) / 2)) +
     (commute.40.to.44 * ((40 + 44) / 2)) +
     (commute.45.to.59 * ((45 + 59) / 2)) +
     (commute.60.to.89 * ((60 + 89) / 2)) +
     (commute.more.90 * 120)
   ) /
  workers.not.at.home

head(avg.commute)
```

## Create new feature median household income

```{r}
median.household.income <- df$Households..Median.household.income.in.1999
head(median.household.income)
```

## Create new feature fraction of houses that are owner occupied

```{r}
occupied.housing.units <- df$Occupied.housing.units..Total
owner.occupied.units <- df$Occupied.housing.units..Owner.occupied

fraction.owner.occupied <- owner.occupied.units / occupied.housing.units
head(fraction.owner.occupied)
```

## Create output file

```{r}
r <- data.frame(census.tract=census.tract,
                avg.commute=avg.commute,
                median.household.income=median.household.income,
                fraction.owner.occupied=fraction.owner.occupied)

# drop records with NaN values in the newly-created variables
missing.commute <- is.na(r$avg.commute)
missing.income <- is.na(r$median.household.income)
missing.occupied <- is.na(r$fraction.owner.occupied)

cat("# of missing commute", sum(missing.commute), "\n")
cat("# of missing income", sum(missing.income), "\n")
cat("# of missing occupied", sum(missing.occupied), "\n")

some.missing <- missing.commute | missing.income | missing.occupied
cat("# with some missing", sum(some.missing), "\n")

a <- r[!some.missing,]  # drop observations with any missing value
cat("# obs retained", nrow(a))

write.table(a,
            path.census.out,
            row.names=FALSE,
            quote=FALSE,
            sep="\t")  # same separator as original file

head(a)
```

