Analysis of Raw Deeds File
========================================================

This R Markdown document performs an analysis of the 8 raw deeds files.

Settings Control Execution
--------------------------

Turn warnings into errors
```{r}
options(warn=2)
```

These settings control execution.

```{r}
settings <- list(testing=FALSE
                 ,dir.laufer="../data/raw/from-laufer-2010-05-11"
                 ,debugging=TRUE
                 )
```

The working directory is `r getwd()`.

Read Input Files
-----------

The input files come from Steve Laufer. Each has a suffix number that
identifies the file. The files are compressed. Only certain fields are 
analyzed, so the remaining are dropped

```{r}
ReadOne <- function(num) {
  path <- paste(settings$dir.laufer,
                "/deeds/CAC06037F",
                num,
                ".txt.gz",
                sep="")
  df <- read.table(path, header=TRUE, sep="\t", quote="", comment.char="", na.strings="",
                   stringsAsFactors=FALSE,
                   encoding="UTF-8",
                   nrows=ifelse(settings$testing,1000,-1))
  if (FALSE & settings$debugging) {
    cat("num", num, "\n")
    print(names(df))
    str(df$APN.UNFORMATTED)
    str(df$APN.FORMATTED)
    str(df$ORIGINAL.APN)  # not present
    str(df$SALE.DATE)
    str(df$RECORDING.DATE)
    str(df$DOCUMENT.TYPE.CODE)
    str(df$TRANSACTION.TYPE.CODE)
    str(df$MULTI.APN.FLAG.CODE)
    str(df$MULTI.APN.COUNT)
    str(df$PRI.CAT.CODE)
    summary(df)
  }
  if (length(df$ORIGINAL.APN) != 0)
    stop("found original apn field")
  if (length(df$LATITUDE) != 0)
    stop("found latitude field")
  if (length(df$LONGITUDE) != 0)
    stop("found longitude field")
  r <- data.frame(APN.UNFORMATTED=df$APN.UNFORMATTED,
                  APN.FORMATTED=df$APN.FORMATTED,
                  APN.SEQUENCE.NUMBER=df$APN.SEQUENCE.NUMBER,
                  SALE.AMOUNT=df$SALE.AMOUNT,
                  SALE.CODE=df$SALE.CODE,
                  SALE.DATE=df$SALE.DATE,
                  RECORDING.DATE=df$RECORDING.DATE,
                  SALE.CODE=df$SALE.CODE,
                  DOCUMENT.TYPE.CODE=df$DOCUMENT.TYPE.CODE,
                  TRANSACTION.TYPE.CODE=df$TRANSACTION.TYPE.CODE,
                  MULTI.APN.FLAG.CODE=df$MULTI.APN.FLAG.CODE,
                  MULTI.APN.COUNT=df$MULTI.APN.COUNT,
                  PRI.CAT.CODE=df$PRI.CAT.CODE)
  cat(sprintf("read %d records from file %d\n", nrow(r), num))
  r
}

ReadAll <- function() {
  r <- rbind(ReadOne(1),
             ReadOne(2),
             ReadOne(3),
             ReadOne(4),
             ReadOne(5),
             ReadOne(6),
             ReadOne(7),
             ReadOne(8))
}

deeds <- ReadAll()
```

In file 5, data record 945 has an NA value for APN.FORMATTED. Replace it, otherwise the factors
don't merge. Also, many other records in file 5 have this problem.

```{r}
r5 <- ReadOne(5)
deeds <- rbind(ReadOne(1), ReadOne(2), ReadOne(3), ReadOne(4))
str(r5)
str(r5$APN.FORMATTED)
sum(is.na(r5$APN.FORMATTED))
```

Summarize
-----------

Note that read.table's stringsAsFactors=FALSE is not honored.

```{r}
str(deeds)
summary(deeds)
```

APNs
---------

Deeds has `r nrow(deeds)` observations.

First few APNs

```{r}
deeds$APN.UNFORMATTED[1:10]
deeds$APN.FORMATTED[1:10]
```

Both unformatted and formatted APNs are often not present

field|num present|num missing
-----|-----------|-----------
APN.FORMATTED|`r sum(!is.na(deeds$APN.FORMATTED))` | `r sum(is.na(deeds$APN.FORMATTED))`
APN.UNFORMATTED|`r sum(!is.na(deeds$APN.UNFORMATTED))` | `r sum(is.na(deeds$APN.UNFORMATTED))`

Determine whether the APN fields can be converted to an integer.
```{r}
only.digits <- "^[0123456789]+$"
apn.formatted.is.numeric <- grepl(only.digits, gsub("-", "", deeds$APN.FORMATTED))
num.numeric.formatted <- sum(apn.formatted.is.numeric)
apn.unformatted.is.numeric <- grepl(only.digits, deeds$APN.UNFORMATTED)
num.numeric.unformatted <- sum(apn.unformatted.is.numeric)
```
`r num.numeric.formatted`  of the formatted APNs convert to integers.

`r num.numeric.unformatted` of the unformatted APNs convert to integers. 



APN SEQUENCE NUMBERs
--------------------

The sequence numbers are all in {0, 1}.
```{r}
table(deeds$APN.SEQUENCE.NUMBER, useNA="ifany")
```

Price
------------

Whas price for the entire parcel?

Code|Meaning
----|------
C|CONFIRMED
E|ESTIMATED
F|SALE PRICE (FULL)
L|SALE PRICE (PARTIAL)
N|NOT OF PUBLIC RECORD
P|SALE PRICE (PARTIAL)
R|LEASE
U|UNKNOWN
V|VERIFIED

```{r}
table(deeds$SALE.CODE, useNA="ifany")
```

Distribution of sale amounts, when price is in full.

```{r}
full.price.indices <- which(deeds$SALE.CODE=="F")
full.prices <- deeds$SALE.AMOUNT[full.price.indices]
hist(full.prices, breaks=100, col="blue")
```

Distribution of non-zero sales prices
```{r}
hist(full.prices[which(full.prices != 0)], breaks=100, col="blue")
```

Boxplot of sale amounts, when price is in full.
```{r}
boxplot(full.prices, col="blue")
```

Prices over time should increase.
```{r}
year <- (deeds$RECORDING.DATE / 10000)[full.price.indices]
plot(year, full.prices, pch=19, col="blue")
```






Transaction Dates
-----------------

The sale date is missing `r sum(is.na(deeds$SALE.DATE))` times.

The recording date is missing `r sum(is.na(deeds$RECORDING.DATE))` times.

Document Types
--------------

Code|Meaning
----|-------
C|CONSTRUCTION LOAN
CD|CORRECTION DEED
F|FINAL JUDGEMENT
G|GRANT DEED
L|LIS PENDENS - NON CALIFORNIA
L|LIENS (STATEWIDE CA)
N|NOTICE OF DEFAULT
Q|QUIT CLAIM
R|RELEASE
S|LOAN ASSIGNMENT 
T|DEED OF TRUST
U|FORECLOSURE
X|MULTI CNTY/ST OR OPEN-END MORTGAGE
Z|NOMINAL


Most of the deeds are Grant(G) deeds. Trust deeds (T) are present.
```{r}
table(deeds$DOCUMENT.TYPE.CODE, useNA="ifany")
```

Transaction Types
-----------------

Meaning of codes (from CoreLogic's documentation file):

Code | Meaning
-----|--------
001|RESALE
002|REFINANCE
003|SUBDIVISION/NEW CONSTRUCTION
004|TIMESHARE
006|CONSTRUCTION LOAN 
007|SELLER CARRYBACK
009|NOMINAL

```{r}
table(deeds$TRANSACTION.TYPE.CODE, useNA="ifany")
```

Sale Code
---------

Sale code is the financial consideration.

Code|Meaning
----|------
C|CONFIRMED
E|ESTIMATED
F|SALE PRICE (FULL)
L|SALE PRICE (PARTIAL)
N|NOT OF PUBLIC RECORD
P|SALE PRICE (PARTIAL)
R|LEASE
U|UNKNOWN
V|VERIFIED

Distribution of sale codes
```{r}
table(deeds$SALE.CODE, useNA="ifany")
```



Multi APNs on One Deed
----------------------

Distribution of number of parcels in the deed:
```{r}
table(deeds$MULTI.APN.COUNT, useNA="ifany")
```

I'm going to assume that 0 in the count means 1.
```{r}
assumed.1.parcel <- deeds$MULTI.APN.COUNT == 0 | deeds$MULTI.APN.COUNT == 1
```

That assumption means that I will say `r sum(assumed.1.parcel)` deeds are for 1 parcel.

Code|Meaning
----|---
D|MULTI / DETAIL PARCEL SALE
M|MULTIPLE PARCEL SALE
S|SPLIT PARCEL SALE

Instead of assuming the 0 count is actually 1, I could use the MULTI APN FLAG CODE.

Distribution of MULTI.APN.FLAG.CODE
```{r}
table(deeds$MULTI.APN.FLAG.CODE, useNA="ifany")
```

Conclusion: Use the MULTI APN FLAG CODE would result in a slightly different count. (Examine
the NA field in the above table.)

Arms-Length Sale
-----------------

Code|Meaning
----|-----
A|ARMS LENGTH TRANSACTION
B|NON ARMS LENGTH - PURCHASE
C|NON ARMS LENGTH - NON PURCHASE
D|NON PURCHASE
E|TIMESHARE
F|NOTICE OF DEFAULT
G|ASSIGNMENT
H|RELEASE

Distribution of PRI.CAT.CODE
```{r}
table(deeds$PRI.CAT.CODE, useNA="ifany")
```

Many non-arms-length deeds are present.







