# deeds-al.Rmd

Create the output file holding certain features of the arms-length deeds. 

I would have retained all features, but my system has only 15 GB RAM, not enough
to keep all the features using this programmatic approach. Instead, could have
created 8 files with just certain deeds and merged them all.

Retain only features of deeds, not features of the parcels. The idea is to
rely on the parcels file exclusive for parcels features.

An earlier versions dropped features that were all missing and always had the
same value. This was an attempt to remove non-informative features, but it
complicates my ability to understand what the programs are doing.

An earlier version also attempted to remove duplicated deeds, but that code was 
broken. Now all arms-length deeds are passed on.

Files read and the deeds files the Laufer directory. This directory contains
the file Steve Laufer used in his experiments.

Record layout for the input is in 1080_Record_layout.csv

```{r}
laufer.dir="../data/raw/from-laufer-2010-05-11"

```

Files written.

```{r}
dir.output="../data/v6/output/"
path.deeds.out <- paste(dir.output,
                         "deeds-al.csv",
                         sep="")
```

## Control the script

Set script control variables.

```{r}
testing <- FALSE
```

Initialize R.

```{r}
options(warn=2)  # turn warnings into errors
set.seed(1)      # random number generator seed
require("compiler")
enableJIT(3)     # 3 ==> maximum JIT level
```

Source files here, now that the JIT level is set

```{r}
source("RemoveNonInformative.R")
```

## Define function to read a deeds file

Read one deeds file with suffix "num" in its file name.

```{r}
ReadDeedsFile <- function(num) {
  # Read the deeds file containing num in its name
  # Args:
  # num: number of input file; in {1, 2, ..., 8}
  #
  # Value: a list
  # $df : data.frame containing selected features for arms-length deeds
  # $num.dropped : number of non-arms-length deeds dropped from $df

  # read the deeds file
  # Note: In file 5, data record 945 has an NA value for APN.FORMATTED
  path <- paste(laufer.dir,
                "/deeds/CAC06037F",
                num,
                ".txt.gz",
                sep="")
  cat("\nreading deeds file", path, "\n")
  # NOTE: Don't convert strings to factors, because the other input
    # files may have different values than this file
  df <- read.table(path,
                   header=TRUE,
                   sep="\t",
                   quote="",
                   comment.char="",
                   stringsAsFactors=FALSE,
                   na.strings="",
                   nrows=ifelse(testing,1000,-1))

  cat("records in", path, nrow(df), "\n")

  # keep the features we want
  # add features to trace back to original source
  r <- data.frame(
    APN.UNFORMATTED=df$APN.UNFORMATTED, 
    APN.FORMATTED=df$APN.FORMATTED,
    SALE.AMOUNT=df$SALE.AMOUNT,
    SALE.DATE=df$SALE.DATE,
    RECORDING.DATE=df$RECORDING.DATE,
    DOCUMENT.TYPE.CODE=df$DOCUMENT.TYPE.CODE,
    TRANSACTION.TYPE.CODE=df$TRANSACTION.TYPE.CODE,
    SALE.CODE=df$SALE.CODE,
    MULTI.APN.FLAG.CODE=df$MULTI.APN.FLAG.CODE,
    MULTI.APN.COUNT=df$MULTI.APN.COUNT,
    PRIOR.SALES.DATE=df$PRIOR.SALES.DATE,
    PRIOR.SALES.AMOUNT=df$PRIOR.SALES.AMOUNT,
    PRIOR.MULTI.APN.FLAG.CODE=df$PRIOR.MULTI.APN.FLAG.CODE,
    PRIOR.MULTI.APN.COUNT=df$PRIOR.MULTI.APN.COUNT,
    PRI.CAT.CODE=df$PRI.CAT.CODE,
    RESALE.NEW.CONSTRUCTION.CODE=df$RESALE.NEW.CONSTRUCTION.CODE,
    deed.file.number=rep(num, nrow(df)),
    deed.record.number=1:nrow(df))


  # keep only arms-length deeds and deeds with numeric apns
  arms.length <- r$PRI.CAT.CODE == "A"  # arms-length deeds
  r <- r[arms.length, ]
  
  list(df=r, num.dropped=nrow(r) - sum(arms.length))
  }
```

## Read the deed files

Concatenate all the deeds files into one data frame.

```{r}
ReadAll <- function() {
  # RETURN: list
  # $df : data frame containing only arms-length deeds and certain features
  # $num.dropped : number of non arms-length deeds found and dropped
  df <- NULL
  num.dropped <- 0
  for (file.number in 1:8) {
    a.list<- ReadDeedsFile(file.number)
    df = rbind(df, a.list$df)
    num.dropped <- num.dropped + a.list$num.dropped
  }
  list(df = df, num.dropped = num.dropped)
}

all <- ReadAll()
cat('total number of non-arms length deeds', all$num.dropped, '\n')
str(all$df)
summary(all$df)
```
## Write deeds to csv

```{r}
cat("writing arms-length deeds to", path.deeds.out, "\n")
write.csv(all,
          file=path.deeds.out,
          quote=FALSE)
```
