# parcels-sfr.Rmd

Create the output file holding the single-family-residential parcels. 

In order to cope with having only 16 GB of RAM on my system, only features 
needed downstream in the pipeline are retained.

Files read and the deeds files the Laufer directory. This directory contains
the file Steve Laufer used in his experiments.

```{r}
laufer.dir="../data/raw/from-laufer-2010-05-11"

```

Files written.

```{r}
dir.output="../data/v6/output/"
path.parcels.out <- paste(dir.output,
                          "parcels-sfr.csv",
                          sep="")
```

## Control the script

Set script control variables.

```{r}
testing <- FALSE
```

Initialize R.

```{r}
options(warn=2)  # turn warnings into errors
set.seed(1)      # random number generator seed
require("compiler")
enableJIT(3)     # 3 ==> maximum JIT level
```

Source files here, now that the JIT level is set

```{r}
source("RemoveNonInformative.R")
```

## Define function to read a taxroll file

The taxroll files contain the parcels.

Define a global variable to keep track of number of non-sfr parcels.
```{r}
num.not.sfr <- 0
```

Define function.

```{r}
ReadParcelsFile <- function(num) {
  path <- paste(laufer.dir,
                "/tax/CAC06037F",
                num,
                ".txt.gz",
                sep="")
  cat("reading tax file", path, "\n")
  df <- read.table(path,
                   header=TRUE,
                   sep="\t",
                   quote="",
                   comment="",
                   stringsAsFactors=FALSE,
                   na.strings="",
                   nrows=ifelse(testing,1000,-1))
  cat("records in", path, nrow(df), "\n")
  
  # select the fields we want
  r <-
    data.frame(APN.UNFORMATTED=df$APN.UNFORMATTED,
               APN.FORMATTED=df$APN.FORMATTED,
               CENSUS.TRACT=df$CENSUS.TRACT,
               ZONING=df$ZONING,
               UNIVERSAL.LAND.USE.CODE=df$UNIVERSAL.LAND.USE.CODE,
               VIEW=df$VIEW,
               LOCATION.INFLUENCE.CODE=df$LOCATION.INFLUENCE.CODE,
               NUMBER.OF.BUILDINGS=df$NUMBER.OF.BUILDINGS,
               PROPERTY.CITY=df$PROPERTY.CITY,
               PROPERTY.ZIPCODE=df$PROPERTY.ZIPCODE,
               TOTAL.VALUE.CALCULATED=df$TOTAL.VALUE.CALCULATED,
               LAND.VALUE.CALCULATED=df$LAND.VALUE.CALCULATED,
               IMPROVEMENT.VALUE.CALCULATED=df$IMPROVEMENT.VALUE.CALCULATED,
               TAX.YEAR=df$TAX.YEAR,
               LAND.SQUARE.FOOTAGE=df$LAND.SQUARE.FOOTAGE,
               UNIVERSAL.BUILDING.SQUARE.FEET=df$UNIVERSAL.BUILDING.SQUARE.FEET,
               LIVING.SQUARE.FEET=df$LIVING.SQUARE.FEET,
               YEAR.BUILT=df$YEAR.BUILT,
               EFFECTIVE.YEAR.BUILT=df$EFFECTIVE.YEAR.BUILT,
               BEDROOMS=df$BEDROOM,
               TOTAL.ROOMS=df$TOTAL.ROOMS,
               TOTAL.BATHS.CALCULATED=df$TOTAL.BATHS.CALCULATED,
               AIR.CONDITIONING.CODE=df$AIR.CONDITIONING.CODE,
               BASEMENT.FINISH.CODE=df$BASEMENT.FINISH.CODE,
               BLDG.CODE=df$BLDG.CODE,
               BLDG.IMPROVEMENT.CODE=df$BLDG.IMPROVEMENT.CODE,
               CONDITION.CODE=df$CONDITION.CODE,
               CONSTRUCTION.TYPE.CODE=df$CONSTRUCTION.TYPE.CODE,
               EXTERIOR.WALLS.CODE=df$EXTERIOR.WALLS.CODE,
               FIREPLACE.INDICATOR.FLAG=df$FIREPLACE.INDICATOR.FLAG,
               FIREPLACE.NUMBER=df$FIREPLACE.NUMBER,
               FIREPLACE.TYPE.CODE=df$FIREPLACE.TYPE.CODE,
               FOUNDATION.CODE=df$FOUNDATION.CODE,
               FLOOR.CODE=df$FLOOR.CODE,
               FRAME.CODE=df$FRAME.CODE,
               GARAGE.CODE=df$GARAGE.CODE,
               HEATING.CODE=df$HEATING.CODE,
               MOBILE.HOME.INDICATOR.FLAG=df$MOBILE.HOME.INDICATOR.FLAG,
               PARKING.SPACES=df$PARKING.SPACE,
               PARKING.TYPE.CODE=df$PARKING.TYPE.CODE,
               POOL.FLAG=df$POOL.FLAG,
               POOL.CODE=df$POOL.CODE,
               QUALITY.CODE=df$QUALITY.CODE,
               ROOF.COVER.CODE=df$ROOF.COVER.CODE,
               ROOF.TYPE.CODE=df$ROOF.TYPE.CODE,
               STORIES.CODE=df$STORIES.CODE,
               STYLE.CODE=df$STYLE.CODE,
               UNITS.NUMBER=df$UNITS.NUMBER,
               ELECTRIC.ENERGY.CODE=df$ELECTRIC.ENERGY.CODE,
               FUEL.CODE=df$FUEL.CODE,
               SEWER.CODE=df$SEWER.CODE,
               WATER.CODE=df$WATER.CODE,
               parcel.file.number=rep(num,nrow(df)),
               parcel.record.number=1:nrow(df))
  df <- NULL  # try to return RAM
  
  # keep only single-family residence parcels
  # CHECK FOR NAs
  sfr <- r$UNIVERSAL.LAND.USE.CODE == 163
  not.sfr <- sum(!sfr)
  num.not.sfr <<- num.not.sfr + not.sfr
  cat(sprintf("dropping %d parcels that are not SFR\n",
              not.sfr))
  r <- r[sfr, ]
  cat("number of SFR parcels", nrow(r), "\n")
  r
}
```

## Read the parcels

Concatenate them into one big data frame.

```{r}
df.all <- rbind(ReadParcelsFile(1),
                ReadParcelsFile(2),
                ReadParcelsFile(3),
                ReadParcelsFile(4),
                ReadParcelsFile(5),
                ReadParcelsFile(6),
                ReadParcelsFile(7),
                ReadParcelsFile(8))

  cat("number of SFR parcels retained", nrow(df.all), "\n")
  cat(sprintf("across all parcels files, dropped %d that were not SFR\n",
              num.not.sfr))
```

## Drop features that are non-informative

```{r}
df.all <- RemoveNonInformative(df.all, cat)
cat("parcels columns retained (others dropped because non informative)\n")
for (colname in colnames(df.all)) {
  cat(" ", colname, "\n")
}
```      
  
## Remove duplicate parcels

```{r}
  nondups <- unique(df.all)
  cat(sprintf("dropped %d parcels that were duplicates of others\n",
              nrow(df.all) - nrow(nondups)))
  cat("number of unique relevant parcels", nrow(nondups), "\n")

  # try to return some memory to the OS
  df.all <- NULL
  gc()
```

## Write unique parcels to CSV

```{r}
cat("writing unique single-family parcels to",
    path.parcels.out, "\n")
write.csv(nondups,
          file=path.parcels.out,
          quote=FALSE)
cat("finished\n")
```


