Program Simple Model.R
========================================================

Create and test some simple models on the future prediction problem.

Simple models include
- linear regression
- weighted linear regression
- linear regression by neighborhood (zip code, zip +4, census tract)
- CART
- random forest
- houses near open space
- houses near shopping (use tax roll to identify shopping)
- houses near factories

Q: can we do better than Sumit? He used 11 months to predict the next 1 month.

Future prediction means to predict the near future (1 month) given all history
up to that time. For now, don't test retrogression.

Create and test 3 different linear regression models:
- predict last 10% of 2004 using just data from first 90% of 2004
- predict last 10% of 2004 using all prior data
- predict last 10% of 2004 using all prior and future data

Run this document by opening it in RStudio and clicking the Knit HTML button 
in the RStudio header.

# Initialize

Set control variables.

```{r}
control <- list(
  dir.output = '../data/v6/output/',
  testing <- TRUE,
  nrows = 1000)
control$path.in = paste0(control$dir.output, "transactions-subset1.csv")
```

Initialize R.

```{r}
options(warn=2)  # turn warnings into errors
set.seed(1)      # random number generator seed
require("compiler")
enableJIT(3)     # 3 ==> maximum JIT level
```
Source files here. Nothing to source for now.
```{r}
source('Center.R')
source('ReadAndTransformTransactions.R')
source('SplitDate.R')
```
# Read data
```{r}
df <- ReadAndTransformTransactions(control$path.in, control$nrows, TRUE)
str(df)
summary(df)
```
# Subset selection

Return logical vectors that indicate the selected features, which are the
training data.

Select transactions with certain dates
```{r}
Is2004JanNov <- function(df) {
  (df$year == 2004) & (df$month != 12)
}

Is2004Dec <- function(df) {
  (df$year == 2004) & (df$month == 12)
}

test.all <- function() {
  df <- data.frame(year <- c(rep(2004, 12), rep(2005, 12)),
                   month <- c(1:12, 1:12))
  result <- Is2004JanNov(df)
  expected <- c(rep(TRUE, 11), FALSE, rep(FALSE, 12))
  stopifnot(result == expected)
  result <- Is2004Dec(df)
  expected <- c(rep(FALSE, 11), TRUE, rep(FALSE, 12))
  stopifnot(result == expected)
}
test.all()
```
# Define experiments
## Experiment 1
Linear regression unweighted, train on first 11 months of 2004, test on Dec
2004. Use factors Chopra used except for those we don't have, namely prior sale
price (we could create this) and quality of the school.
```{r}
Ex1 <- function(data) {
  cat('Ex1\n')
  # RETURN fraction of estimates within 10 percent.
  formula <- log.price ~ 
    centered.log.living.area +
    centered.year.built +
    centered.log.bedrooms +
    centered.log.bathrooms +
    centered.log1p.parking.spaces +
    centered.log.land.value +
    centered.log.improvement.value +
    centered.fraction.improvement.value +
    factor.parking.type +
    factor.has.pool +
    factor.is.new.construction +
    factor.foundation.type +
    factor.roof.type +
    factor.site.influence +
    centered.latitude +
    centered.longitude +
    centered.log.median.household.income +
    centered.fraction.owner.occupied +
    centered.avg.commute.time
    
  m <- lm(formula = formula,
          data = data,
          subset = Is2004JanNov(data))
  str(fitted)
  summary(fitted)
  predictions <- predict(m, newdata=data[Is2004Dec(data)])
  stop('predict 2004Dec and determine fraction within 10%')
  
}
```
# Experiment 2
Add or remove variables from experiment 1. See page 281 in cookbook.

# Run selected experiments
```{r}
Ex1(df)
```

