# transactions-subset1-to-impute.Rmd

Determine features in subset 1 that are to be imputed.

Analyze the qualitative features in subset 1 file to provide information 
to write the report and to determine which features should be imputed.

Files read and written:

```{r}
input.file <- "transactions-subset1.csv"
table1.file <- "transactions-subset1-analyze-qualitative-features.tex"
table2.file <- "transactions-subset1-analyze-imputed-features.tex"
imputed.file <- "transactions-subset1-to-impute.csv"
```

Output files created:
* `r table1.file`: table for report showing all qualitative features 
* `r imputed.file`: csv containing names of features that should be imputed

## Control the script

Set script control variables.

```{r}
testing <- FALSE
output.dir <- "../data/v6/output/"
script.name <- "createSubset1Imputed"
path.input.file <- paste(output.dir, input.file, sep="")
path.table1.file <- paste(output.dir, table1.file, sep="")
path.table2.file <- paste(output.dir, table2.file, sep="")
path.imputed.file <- paste(output.dir, imputed.file, sep="")
```

Initialize R.
```{r}
options(warn=2)  # turn warnings into errors
set.seed(1)      # random number generator seed
require("compiler")
enableJIT(3)     # 3 ==> maximum JIT level
```

Source files here, now that the JIT level is set
```{r}
# nothing to source for now
```

## Read the input

Read input transactions into dataframe `t`.

```{r}
path.input.file
t <- read.csv(path.input.file,
              nrows=ifelse(testing, 1000, -1))

str(t)
summary(t)
```

## Summary of qualitative features (table 1)

This code generates table1.

````{r}
table1Lines <- list()

# These function accumulate their results into the global table1Lines
# The goal is to write the table to a txt file that is imported into the latex
# documentation file. That way I don't have to copy and paste results.
# The table in the txt file comes into latex as a verbatim block. Maybe later
# it can be turned into a latex table with fancy formatting.

Header1 <- function() {
  # Write the header
  table1Lines <<- list("\\begin{verbatim}")
  table1Lines <<- c(table1Lines, "NUMBER OF UNIQUES CODES AND NUMBER OF MISSING CODES")
  table1Lines <<- c(table1Lines, "Number of Uniques Includes Any NAs")
  table1Lines <<- c(table1Lines, " ")
  line <- sprintf("%40s %11s %11s %12s", 
                  "Qualitative Feature", "Num Uniques", "Num NAs", "Fraction NAs")
  table1Lines <<- c(table1Lines, line)
}

Detail1 <- function(columnName) {
  # Write a detail line
  values <- t[[columnName]]
  #str(values)
  nUniques <- length(unique(values))
  nNAs <- sum(is.na(values))
  fractionNAs <- nNAs / nrow(t)
  line <- sprintf("%40s %11d %11d %12.4f", columnName, nUniques, nNAs, fractionNAs)  
  invisible(table1Lines <<- c(table1Lines, line))

  #Data1(columnName, nUniques, nNAs, fractionNAs)
  #paste(columnName, "|", nUniques, "|", nNAs, "|", fractionNAs, sep="")
}

Footer1 <- function() {
  # Write the footer
  table1Lines <<- c(table1Lines, "\\end{verbatim}")
}
```

Now generate table 1.

```{r}
Header1()
Detail1("RESALE.NEW.CONSTRUCTION.CODE")
Detail1("ZONING")
Detail1("VIEW")
Detail1("LOCATION.INFLUENCE.CODE")
Detail1("AIR.CONDITIONING.CODE")
Detail1("CONDITION.CODE")
Detail1("CONSTRUCTION.TYPE.CODE")
Detail1("EXTERIOR.WALLS.CODE")
Detail1("FIREPLACE.TYPE.CODE")
Detail1("FOUNDATION.CODE")
Detail1("FLOOR.CODE")
Detail1("FRAME.CODE")
Detail1("GARAGE.CODE")
Detail1("HEATING.CODE")
Detail1("PARKING.TYPE.CODE")
Detail1("POOL.FLAG")
Detail1("POOL.CODE")
Detail1("QUALITY.CODE")
Detail1("ROOF.COVER.CODE")
Detail1("ROOF.TYPE.CODE")
Detail1("STYLE.CODE")
Detail1("SEWER.CODE")
Detail1("WATER.CODE")
Footer1() 
```

Print table 1 into the HMTL document for ease of reviewing.
```{r}
str(table1Lines)
```

Write table1 to external file.

```{r}
writeLines(as.character(table1Lines), con=path.table1.file)
```

## Details on Qualitative Fields

Consider each feature then decide what to do with it
* Drop fields that cannot be used. Delete these features from the `t` data frame.
* Recode fields that need to be fixed up. Replace original variable with new one.
* Remember names of features that will be imputed in a later step.
* Sometime both recoded the field and remember it for later imputation.

Keep track of names of features for the imputation.
```{r}
imputedFeatures <- list()
```

### RESALE NEW CONSTRUCTION CODE retained

```{r}
table(t$RESALE.NEW.CONSTRUCTION.CODE, useNA="ifany")
```

### ZONING imputed

```{r}
#table(t$ZONING, useNA="ifany")   DROP, because the long print out clutters the HTML file
imputedFeatures <- c(imputedFeatures, "ZONING")
```

### VIEW droped

```{r}
table(t$VIEW, useNA="ifany")
t$VIEW <- NULL
```

### LOCATION INFLUENCE CODE recoded

```{r}
table(t$LOCATION.INFLUENCE.CODE, useNA="ifany")
t$LOCATION.INFLUENCE.CODE.recoded <- ifelse(is.na(t$LOCATION.INFLUENCE.CODE),
                                            "NOT SPECIAL",
                                            t$LOCATION.INFLUENCE.CODE)
t$LOCATION.INFLUENCE.CODE <- NULL
table(t$LOCATION.INFLUENCE.CODE.recoded, useNA="ifany")
```

### AIR CONDITIONING CODE recoded

```{r}
table(t$AIR.CONDITIONING.CODE, useNA="ifany")
# set missing values to 000 which is the code for non
t$AIR.CONDITIONING.CODE.recoded <- ifelse(is.na(t$AIR.CONDITIONING.CODE),
                                          "000",
                                          t$AIR.CONDITIONING.CODE)
t$AIR.CONDITIONING.CODE <- NULL
```

### CONDITION CODE imputed

```{r}
table(t$CONDITION.CODE, useNA="ifany")
imputedFeatures <- c(imputedFeatures, "CONDITION.CODE")
```

### CONSTRUCTION TYPE CODE imputed

```{r}
table(t$CONSTRUCTION.TYPE.CODE, useNA="ifany")
imputedFeatures <- c(imputedFeatures, "CONSTRUCTION.TYPE.CODE")
```

### EXTERIOR WALLS CODE imputed

```{r}
table(t$EXTERIOR.WALLS.CODE, useNA="ifany") # to be imputed
imputedFeatures <- c(imputedFeatures, "EXTERIOR.WALLS.CODE")
```

### FIREPLACE TYPE CODE dropped

```{r}
table(t$FIREPLACE.TYPE.CODE, useNA="ifany")
t$FIREPLACE.TYPE.CODE <- NULL
```
### FOUNDATION CODE recoded imputed

```{r}
table(t$FOUNDATION.CODE, useNA="ifany")
t$FOUNDATION.CODE.recoded <- ifelse(t$FOUNDATION.CODE == "001",
                                    NA,
                                    t$FOUNDATION.CODE)
t$FOUNDATION.CODE <- NULL
table(t$FOUNDATION.CODE.recoded, useNA="ifany")
imputedFeatures <- c(imputedFeatures, "FOUNDATION.CODE.recoded")
```

### FLOOR CODE dropped

```{r}
table(t$FLOOR.CODE, useNA="ifany")
t$FLOOR.CODE <- NULL
```

### FRAME CODE dropped

```{r}
table(t$FRAME.CODE, useNA="ifany")
t$FRAME.CODE <- NULL
```

### GARAGE CODE recoded

```{r}
table(t$GARAGE.CODE, useNA="ifany")
# set missing values to 000 which is the code for no garage
t$GARAGE.CODE.recoded <- ifelse(is.na(t$GARAGE.CODE),
                                "000",
                                t$GARAGE.CODE)
t$GARAGE.CODE <- NULL
```

### HEATING CODE imputed

```{r}
table(t$HEATING.CODE, useNA="ifany")
imputedFeatures <- c(imputedFeatures, "HEATING.CODE")
```

### PARKING TYPE CODE recoded imputed

```{r}
# When parking type is missing, often there are zero parking spaces
table(t$PARKING.TYPE.CODE, useNA="ifany")
table(t$PARKING.SPACES, useNA="ifany")
table(t$PARKING.TYPE.CODE, t$PARKING.SPACES, useNA="ifany")
# set missing PARKING TYPE CODE to 000 (no parking space) iff PARKING SPACE == 0
t$PARKING.TYPE.CODE.recoded <- ifelse(is.na(t$PARKING.TYPE.CODE),
                                      ifelse(t$PARKING.SPACES == 0,
                                             "000", 
                                             NA),
                                      t$PARKING.TYPE.CODE)
imputedFeatures <- c(imputedFeatures, "PARKING.TYPE.CODE.recoded")
```

### POOL FLAG recoded

```{r}
poolFlagTable <- table(t$POOL.FLAG, useNA="ifany")
poolFlagTable
poolFlagTableNAs <- poolFlagTable[length(poolFlagTable)]
poolFlagTableNAs
```

The only valid value is Y
```{r}
t$POOL.FLAG.recoded <- ifelse(is.na(t$POOL.FLAG) | t$POOL.FLAG == "N",
                              "N",
                              "Y")
t$POOL.FLAG <- NULL
```

### POOL CODE recoded

Is POOL CODE present when POOL FLAG recoded is "Y"?
```{r}
poolCodeTable <- table(t$POOL.CODE, useNA="ifany")
poolCodeTable
poolCodeTableNAs <- poolCodeTable[length(poolCodeTable)]
if (poolCodeTableNAs != poolFlagTableNAs)
  stop("assumption about POOL CODE and POOL FLAG NA counts is wrong")
# set missing values to 000 which means no pool
t$POOL.CODE.recoded <- ifelse(is.na(t$POOL.CODE),
                              "000",
                              t$POOL.CODE)
t$POOL.CODE <- NULL
```

### QUALITY CODE imputed

```{r]}
table(t$QUALITY.CODE, useNA="ifany")
imputedFeatures <- c(imputedFeatures, "QUALITY.CODE")
```

### ROOF COVER CODE imputed

```{r}
table(t$ROOF.COVER.CODE, useNA="ifany") 
imputedFeatures <- c(imputedFeatures, "ROOF.COVER.CODE")
```

### ROOF TYPE CODE imputed

```{r}
table(t$ROOF.TYPE.CODE, useNA="ifany")
imputedFeatures <- c(imputedFeatures, "ROOF.TYPE.CODE")
```

### STYLE CODE imputed

```{r}
table(t$STYLE.CODE, useNA="ifany")
imputedFeatures <- c(imputedFeatures, "STYLE.CODE")
```

### SEWER CODE dropped

```{r}
table(t$SEWER.CODE, useNA="ifany")
t$SEWER.CODE <- NULL
```

### WATER CODE imputed

```{r}
table(t$WATER.CODE, useNA="ifany")
imputedFeatures <- c(imputedFeatures, "WATER.CODE")
```



## Qualitative fields containing missing values that are to be imputed

All the fields to be imputed.
```{r}
imputedFeatures
```

Write csv containing names of features to be imputed.
```{r}
cat("featureName\n", file=path.imputed.file)
for (imputedFeature in imputedFeatures) {
  cat(paste(imputedFeature, "\n", sep=""), file=path.imputed.file, append=TRUE)
}
```

Code to create external table of number of NAs in fields to be imputed.

````{r}
table2Lines <- list()  # the lines accumulate into this global variable
runningTotal <- 0

Header2 <- function() {
  # Write the header
  table2Lines <<- list("\\begin{verbatim}")
  table2Lines <<- c(table2Lines, "NUMBER OF MISSING CODES IN FIELDS TO BE IMPUTED")
  table2Lines <<- c(table2Lines, " ")
  line <- sprintf("%40s %11s", "Feature", "Num NAs")
  table2Lines <<- c(table2Lines, line)
}

Detail2 <- function(columnName) {
  # Write a detail line
  values <- t[[columnName]]
  nNAs <- sum(is.na(values))
  runningTotal <<- runningTotal + nNAs
  line <- sprintf("%40s %11d", columnName, nNAs)  
  invisible(table2Lines <<- c(table2Lines, line))
}

Footer2 <- function() {
  # Write the footer
  table2Lines <<- c(table2Lines, "\\end{verbatim}")
}
```

Now generate table 1.

```{r}
Header2()
for (imputedFeature in imputedFeatures) {
  Detail2(imputedFeature)
}
Footer2()
```

Print table 2 into the HMTL document for ease of reviewing.
```{r}
str(table2Lines)
```

Write table2 to external file.

```{r}
writeLines(as.character(table2Lines), con=path.table2.file)
```

Determine 
"finished"