# transactions-al-sfr.Rmd

Create the transactions-al-sfr.csv file containing all transactions for
arms-length deeds for single-family-residential parcels.

Join the deeds and parcels files on the best APN (defined below), not the
formatted or unformatted APN.

Paths to files read.

```{r}
dir.output="../data/v6/output/"
path.census <- paste(dir.output, "census.csv", sep="")
path.deeds <- paste(dir.output, "deeds-al.csv", sep="")
path.parcels <- paste(dir.output, "parcels-sfr.csv", sep="")
path.geocoding <- "../data/raw/geocoding.tsv"
```

Files written.

```{r}
path.out <- paste(dir.output, "transactions-al-sfr.csv", sep="")
```

## Control the script

Set script control variables.

```{r}
testing <- FALSE
```

Initialize R.

```{r}
options(warn=2)  # turn warnings into errors
set.seed(1)      # random number generator seed
require("compiler")
enableJIT(3)     # 3 ==> maximum JIT level
```

Source files here, now that the JIT level is set

```{r}
source("BestApns.R")
source("RemoveNonInformative.R")
```

## Read census file

```{r}
census <- read.table(path.census,
                     header=TRUE,
                     sep="\t",
                     quote="",
                     comment.char="",
                     stringsAsFactors=FALSE,
                     nrows=ifelse(testing, 1000, -1))                  
cat("number of census records read", nrow(census), "\n")
```
## Read deeds file

```{r}
deeds <- read.csv(path.deeds,
                  header=TRUE,
                  quote="",
                  comment="",
                  stringsAsFactors=FALSE,
                  nrows=ifelse(testing, 1000, -1))
cat("number of deeds records read", nrow(deeds), "\n")
```

## Read geocoding

```{r}
geocoding <- read.table(path.geocoding,
                        header=TRUE,
                        sep="\t",
                        quote="",
                        comment="",
                        stringsAsFactor=FALSE,
                        na.strings="",
                        nrows=ifelse(testing, 1000, -1))
cat("number of geocoding records read", nrow(geocoding), "\n")
```

## Read parcels file

```{r}
parcels <- read.csv(path.parcels,
               header=TRUE,
               quote="",
               comment="",
               stringsAsFactors=FALSE,
               nrows=ifelse(testing, 1000, -1))
cat("number of parcels records read", nrow(parcels), "\n")
```

  
## Create the transactions

Find best APNs from the deeds and parcels.

```{r}
deeds$apn.recoded <- BestApns(deeds$APN.UNFORMATTED,
                              deeds$APN.FORMATTED)

# show some examples
deeds$APN.UNFORMATTED[1:100]
deeds$APN.FORMATTED[1:100]
as.character(deeds$apn.recoded)[1:100] 

parcels$apn.recoded <- BestApns(parcels$APN.UNFORMATTED,
                                parcels$APN.FORMATTED)
# show some examples
parcels$APN.UNFORMATTED[1:100]
parcels$APN.FORMATTED[1:100]
as.character(parcels$apn.recoded)[1:100]
```

Merge the deeds and parcels.
```{r}
m1 <- merge(deeds, parcels, by="apn.recoded",
            suffixes=c(".deeds", ".parcels"))
cat("number of deeds and parcels with common recoded.apn",
    nrow(m1),
    "\n")
deeds <- NULL    # try to release RAM
parcels <- NULL
gc()
```

Merge in the census data
```{r}
m2 <- merge(m1, census, by.x="CENSUS.TRACT", by.y = "census.tract")
cat("number of common deeds and parcels with known CENSUS.TRACT",
    nrow(m2),
    "\n")
census <- NULL
gc()
```

Merge in the geocoding data. Note: Some of the latitudes and longitudes are 
zero, an invalid value. Pass these along because some algorithms may be able
to cope with these values.
```{r}
m3 <- merge(m2, geocoding, by.x="apn.recoded", by.y="G.APN")
cat("number of transactions, after considering geocoding",
    nrow(m3),
    "\n")
geocoding <- NULL
gc()
```

Drop some extraneous features.
```{r}
m3$X.deeds <- NULL
m3$APN.UNFORMATTED.deeds <- NULL
m3$APN.FORMATTED.deeds <- NULL
m3$X.parcels <- NULL
m3$APN.UNFORMATTED.parcels <- NULL
```

Write transactions file
```{r}
write.csv(m3, path.out)
cat("number of transactions written:", nrow(m3), "\n")
```

```{r}
cat("Finished\n")
```