# transaction-subset1.Rmd

Create output file transactions-subset1.csv.

Also create two output analysis files:
* transactions-subset1-ranges.tex : ranges of features used to create the subset
* transactions-subset1-excluded.tex: number of observations excluded by each criterion in isolation

The input file is transactions-al-sfr.csv. 

## Initialize

Set control variables.

```{r}
testing <- FALSE
output.dir <- "../data/v6/output/"
path.input <- paste(output.dir, "transactions-al-sfr.csv", sep="")
path.output <- paste(output.dir, "transactions-subset1.csv", sep="")
path.ranges <- paste(output.dir, "transactions-subset1-ranges.tex", sep="")
path.excluded <- paste(output.dir, "transactions-subset1-excluded.tex", sep="")
```

Initialize R.
```{r}
options(warn=2)  # turn warnings into errors
set.seed(1)      # random number generator seed
require("compiler")
enableJIT(3)     # 3 ==> maximum JIT level
```

Source other files here, now that the JIT level is set.

```{r}
# nothing to source for now
```

## Read the transactions into variable t.

```{r}
t <- read.csv(path.input,
              nrows=ifelse(testing, 1000, -1)) 
```

## Summarize the resulting data frame.
```{r}
str(t)
summary(t)
```

## Transaction count
Number of transactions in input file:
```{r}
nrow(t)
```

## Overview of features used to form the subset

Code to generate table of ranges.

```{r}
table1Lines <- list()

# These function accumulate their results into the global table1Lines
# The goal is to write the table to a txt file that is imported into the latex
# documentation file. That way I don't have to copy and paste results.
# The table in the txt file comes into latex as a verbatim block. Maybe later
# it can be turned into a latex table with fancy formatting.

Header1 <- function(feature, min, max, nNAs) {
  table1Lines <<- list("\\begin{verbatim}")
  table1Lines <<- c(table1Lines, "SUMMARY OF FEATURES USED TO GENERATE THE SUBSET")
  table1Lines <<- c(table1Lines, " ")
  line <- sprintf("%40s %10s %10s %10s", "Feature", "Min", "Max", "Num NAs")
  table1Lines <<- c(table1Lines, line)
}

Data1 <- function(feature, min, max, nNAs) {
  line <- sprintf("%40s %10s %10s %10s", feature, min, max, nNAs)  
  #cat("newline", line, "\n")
  invisible(table1Lines <<- c(table1Lines, line))
}

Footer1 <- function() {
  table1Lines <<- c(table1Lines, "\\end{verbatim}")
}

# These functions build the table
Continuous <- function(feature) {
  values <- t[[feature]]
  value.min <- min(values, na.rm=TRUE)
  value.max <- max(values, na.rm=TRUE)
  value.nNAs <- sum(is.na(values))
  Data1(feature, value.min, value.max, value.nNAs)
}

Discrete <- function(feature) {
  values <- t[[feature]]
  value.min <- "-"
  value.max <- "-"
  value.nNAs <- sum(is.na(values))
  Data1(feature, value.min, value.max, value.nNAs)
}
```

Generate the table.

```{r}
Header1()
Continuous("SALE.DATE")
Continuous("RECORDING.DATE")
Continuous("SALE.AMOUNT")
Discrete("DOCUMENT.TYPE.CODE")
Discrete("TRANSACTION.TYPE.CODE")
Discrete("SALE.CODE")
Discrete("MULTI.APN.FLAG.CODE")
Continuous("MULTI.APN.COUNT")
Continuous("NUMBER.OF.BUILDINGS")
Continuous("TOTAL.VALUE.CALCULATED")
Continuous("LAND.VALUE.CALCULATED")
Continuous("IMPROVEMENT.VALUE.CALCULATED")
Continuous("LAND.SQUARE.FOOTAGE")
Continuous("UNIVERSAL.BUILDING.SQUARE.FEET")
Continuous("LIVING.SQUARE.FEET")
Continuous("EFFECTIVE.YEAR.BUILT")
Continuous("BEDROOMS")
Continuous("TOTAL.ROOMS")
Continuous("TOTAL.BATHS.CALCULATED")
Continuous("G.LATITUDE")
Continuous("G.LONGITUDE")
Footer1()
```

Display the generated lines in the HTML file.

```{r}
str(table1Lines)
```

Write lines to external file.
```{r}
writeLines(as.character(table1Lines), con=path.ranges)
```


## Constraint on transaction date

The constraint is that a sale date must be inferable. Often SALE DATE is missing, but RECORDING DATE is always present.
```{r}
sum(is.na(t$SALE.DATE))
sum(is.na(t$RECORDING.DATE))
```

The plan is to infer missing SALE DATE values by subtracting the mean difference between
the SALE DATE and RECORDING DATE for known observations.

```{r}
ToDate <- function(v) {
  # Value: vector of adjusted sales dates
  c <- as.character(v)
  c2 <- gsub("00$", "15", c) # move dates like 1991-01-00 to 1991-01-15
  dates <- as.Date(c2, "%Y%m%d")
  dates
}
```

After adjustment, many of the sales dates are bad.
```{r}
adjusted.sale.date <- ToDate(t$SALE.DATE)
sum(is.na(adjusted.sale.date))
```

All the recording dates adjust to valid dates.
```{r}
adjusted.recording.date <- ToDate(t$RECORDING.DATE)
sum(is.na(adjusted.recording.date))
```

Comparing the raw dates
```{r}
t$SALE.DATE[1:10]
t$RECORDING.DATE[1:10]
```

Comparing the adjusted dates
```{r}
adjusted.sale.date[1:10]
adjusted.recording.date[1:10]
```


The mean differences for known sales dates:
```{r}
adjusted.sale.date.present <- !is.na(adjusted.sale.date)
diff <- adjusted.recording.date[adjusted.sale.date.present] - adjusted.sale.date[adjusted.sale.date.present] 
diff[1:10]
mean.diff <- mean(diff)
mean.diff
````

Create new features to hold the transaction date. Determine number of records without a transaction date.
```{r}
t$transaction.date <- as.Date(ifelse(is.na(adjusted.sale.date),
                                   adjusted.recording.date - mean.diff,
                                   adjusted.sale.date),
                            origin="1970-01-01")
ok.transaction.date <- !is.na(t$transaction.date)
sum(!ok.transaction.date)
```

Comparing all the dates
```{r}
t$SALE.DATE[1:10]
t$RECORDING.DATE[1:10]
adjusted.sale.date[1:10]
adjusted.recording.date[1:10]
transaction.date[1:10]
```


## Constraints on SALE AMOUNT

Range of sale amount includes 0 and some very large values.
```{r}
summary(t$SALE.AMOUNT)
boxplot(t$SALE.AMOUNT, col="blue")
```

The sale amount should not be zero nor over $85 million. The $85 million
amount come from the Wall Street Journal on January 7, 2013, which reported that 
the most expensive Los Angeles house every sold had a price of $85 million.

Number of observations dropped by constraining SALE AMOUNT > 0.
```{r}
ok.sale.amount.positive <- t$SALE.AMOUNT > 0
sum(!ok.sale.amount.positive)
```

Count of number of observerations dropped by constraining SALE AMOUNT , sale.amount.max.cutoff:
```{r}
sale.amount.max.cutoff <- 85e6
sum(!(t$SALE.AMOUNT < sale.amount.max.cutoff))
```

Count number of observations dropped by constaining SALE AMOUNT < $85 million.

```{r}
sum(!(t$SALE.AMOUNT < 85000000))
```

Number of observations dropped by constraining to postive sales amount less than $85 million:
```{r}
ok.sale.amount.not.too.large <- t$SALE.AMOUNT < 85000000
sum(!ok.sale.amount.not.too.large)
```

Define constraint and count excluded obs
```{r}
ok.sale.amount <- t$SALE.AMOUNT > 0 & t$SALE.AMOUNT <= sale.amount.max.cutoff
sum(!ok.sale.amount)
```

## Constraints on DOCUMENT TYPE CODE

The DOCUMENT TYPE CODES is the type of transfer document recorded. The codes are:

Code|Meaning
----|-------
C|CONSTRUCTION LOAN
CD|CORRECTION DEED
F|FINAL JUDGEMENT
G|GRANT DEED
L|LIS PENDENS - NON CALIFORNIA
L|LIENS (STATEWIDE CA)
N|NOTICE OF DEFAULT
Q|QUIT CLAIM
R|RELEASE
S|LOAN ASSIGNMENT 
T|DEED OF TRUST
U|FORECLOSURE
X|MULTI CNTY/ST OR OPEN-END MORTGAGE
Z|NOMINAL

```{r}
table(t$DOCUMENT.TYPE.CODE, useNA="ifany")
```

We want only Grant (G) and Trust (T) deeds.

Number of observations dropped by constraining on Grant and Trust deeds only:
```{r}
ok.document.type.code <- t$DOCUMENT.TYPE.CODE == "G" | t$DOCUMENT.TYPE.CODE == "T"
sum(!ok.document.type.code)
```

## Constraint on TRANSACTION TYPE CODE

This field records the type of transaction. The codes are 

Code | Meaning
-----|--------
001|RESALE
002|REFINANCE
003|SUBDIVISION/NEW CONSTRUCTION
004|TIMESHARE
006|CONSTRUCTION LOAN 
007|SELLER CARRYBACK
009|NOMINAL

```{r}
table(t$TRANSACTION.TYPE.CODE, useNA="ifany")
```

We want only resales of existing properties and new construction sales.

Number of observations dropped by constraining only on type of sale:
```{r}
ok.transaction.type.code <-
  (!is.na(t$TRANSACTION.TYPE.CODE) & 
   (t$TRANSACTION.TYPE.CODE == 1 | t$TRANSACTION.TYPE.CODE == 3))
sum(!ok.transaction.type.code)
```

## Constraint on SALE CODE

Sale code is the financial consideration.

Code|Meaning
----|------
C|CONFIRMED
E|ESTIMATED
F|SALE PRICE (FULL)
L|SALE PRICE (PARTIAL)
N|NOT OF PUBLIC RECORD
P|SALE PRICE (PARTIAL)
R|LEASE
U|UNKNOWN
V|VERIFIED

```{r}
table(t$SALE.CODE, useNA="ifany")
```

We want to use only the deeds that record the full sales price.

Number of observations by constraining only on financial consideration:
```{r}
ok.sale.code <- !is.na(t$SALE.CODE) & t$SALE.CODE == "F"
sum(!ok.sale.code)
```

## Constraint on MULTI APN FLAG CODE and MULTI APN COUNT

We want to retain only transactions that are for one parcel.

Meanings of MULTI APN FLAG CODE.

Code|Meaning
----|---
D|MULTI / DETAIL PARCEL SALE
M|MULTIPLE PARCEL SALE
S|SPLIT PARCEL SALE


MULTI APN FLAG CODE is mostly missing. An undefined X code is also used.
```{r}
table(t$MULTI.APN.FLAG.CODE, useNA="ifany")
```

MULTI APN COUNT if often zero.
```{r}
table(t$MULTI.APN.COUNT, useNA="ifany")
```

I assumed that a one-parcel sale meets these criteria:

- Has NA value in MULTI APN FLAG CODE
- Has a MULTI APN COUNT of 0 or 1

Number of observations dropped by constraining only on one parcel:
```{r}
ok.one.parcel <- is.na(t$MULTI.APN.FLAG.CODE) & t$MULTI.APN.COUNT <= 1
sum(!ok.one.parcel)
```

## Constraint on number of buildings

We should have  one building;
```{r}
table(t$NUMBER.OF.BUILDINGS, useNA="ifany")
ok.number.of.buildings <- t$NUMBER.OF.BUILDINGS == 1
sum(!ok.number.of.buildings)
```

## Constraint on assessed value

The value of the property is estimated by the tax assessor. It's broken down into the land 
value and the improvement value. 

We want no transactions where either is zero or where there sum is zero.
```{r}
summary(t$TOTAL.VALUE.CALCULATED)
summary(t$LAND.VALUE.CALCULATED)
summary(t$IMPROVEMENT.VALUE.CALCULATED)
```



Some values exceed the known maximum of $85,000,000. IS THIS TRUE?
```{r}
summary(t$TOTAL.VALUE.CALCULATED)
summary(t$LAND.VALUE.CALCULATED)
summary(t$IMPROVEMENT.VALUE.CALCULATED)
```

Number of obs eliminated by max value constraint:
```{r}
max.value <- 85000000
assessed.value.not.zero <- 
  t$TOTAL.VALUE.CALCULATED > 0 & 
  t$LAND.VALUE.CALCULATED > 0 & 
  t$IMPROVEMENT.VALUE.CALCULATED > 0
assessed.value.not.too.large <- 
  t$TOTAL.VALUE.CALCULATED < max.value & 
  t$LAND.VALUE.CALCULATED < max.value & 
  t$IMPROVEMENT.VALUE.CALCULATED < max.value
ok.assessed.value <- assessed.value.not.zero & assessed.value.not.too.large
sum(!ok.assessed.value)
```

## Constraint on LAND SQUARE FOOTAGE

Some observations have no land and some have a huge amount of land.

```{r}
summary(t$LAND.SQUARE.FOOTAGE)
```

Quantiles
```{r}
quantile(t$LAND.SQUARE.FOOTAGE, probs=seq(0, 1, .05))
q <- quantile(t$LAND.SQUARE.FOOTAGE, probs=seq(.95, 1, .01))
q
```

Set max land to 99th percentile
```{r}
max.land <- q[5]
max.land
```

Number of observations dropped
```{r}
ok.land.square.footage <- t$LAND.SQUARE.FOOTAGE > 0 & t$LAND.SQUARE.FOOTAGE <= max.land
sum(!ok.land.square.footage)
```


## Constraint on UNIVERSAL BUILDING SQUARE FEET

Some observations have no square feet, some huge:
```{r}
summary(t$UNIVERSAL.BUILDING.SQUARE.FEET)
```

Number of observations dropped by constraining only to some square feet:
```{r}
sum(!(t$UNIVERSAL.BUILDING.SQUARE.FEET > 0))
```

Consider reasonable up to the 99th percentile
```{r}
quantile(t$UNIVERSAL.BUILDING.SQUARE.FEET, probs=seq(0, 1, .05))
q <- quantile(t$UNIVERSAL.BUILDING.SQUARE.FEET, probs=seq(.95, 1, .01))
q
max.building <- q[5]
max.building
```

Number of observations dropped
```{r}
ok.universal.building.square.feet <- 
  t$UNIVERSAL.BUILDING.SQUARE.FEET > 0 & t$UNIVERSAL.BUILDING.SQUARE.FEET <= max.building
sum(!ok.universal.building.square.feet)
```

## Constraint on LIVING SQUARE FEET

Some observations have no square feet, some huge:
```{r}
summary(t$LIVING.SQUARE.FEET)
```

Consider reasonable up to the 99th percentile
```{r}
quantile(t$LIVING.SQUARE.FEET, probs=seq(0, 1, .05))
q <- quantile(t$LIVING.SQUARE.FEET, probs=seq(0.95, 1, .01))
q
max.living <- q[5]
```

Number of observations dropped 
```{r}
ok.living.square.feet <- t$LIVING.SQUARE.FEET > 0 & t$LIVING.SQUARE.FEET <= max.living
sum(!ok.living.square.feet)
```


## Constraints on YEAR BUILT

We use this field during imputation. It should not be zero
.
```{r}
summary(t$YEAR.BUILT)
```

Number of observations dropped only by constraining to some year built:
```{r}
ok.year.built <- t$YEAR.BUILT > 0
sum(!ok.year.built)
```

Distribution of year built
```{r}
table(t$YEAR.BUILT, useNA = "ifany")
```

## Constraints on EFFECTIVE YEAR BUILT

This is the last year of remodeling. It should not be zero.

```{r}
summary(t$EFFECTIVE.YEAR.BUILT)
```

Number of observations dropped only by constraining to some effective year built:
```{r}
ok.effective.year.built <- t$EFFECTIVE.YEAR.BUILT > 0
sum(!ok.effective.year.built)
```

Distribution of EYB
```{r}
table(t$EFFECTIVE.YEAR.BUILT, useNA = "ifany")
```


## Constraint on TOTAL ROOMS

We allow zero bedrooms and zero bathrooms (could be an outhouse), but not zero total rooms.

```{r}
summary(t$BEDROOMS)
summary(t$TOTAL.BATHS.CALCULATED)
summary(t$TOTAL.ROOMS)
```

Percentiles
```{r}
quantile(t$TOTAL.ROOMS, probs=seq(0, 1, 0.05))
q <- quantile(t$TOTAL.ROOMS, probs = seq(.95, 1, .01))
q
max.total.rooms <- q[5]
max.total.rooms
```

Constraint on total rooms
```{r}
ok.total.rooms <- t$TOTAL.ROOMS > 0 & t$TOTAL.ROOMS <= max.total.rooms
sum(!ok.total.rooms)
```

distributions of other rooms conditioned on ok.total.rooms
```{r}
table(t$BEDROOMS[ok.total.rooms], useNA="ifany")
table(t$TOTAL.BATHS.CALCULATED[ok.total.rooms], useNA="ifany")
```
## Constraing on UNITS NUMBER

Number of residential units or apartments.
```{r}
table(t$UNITS.NUMBER, usNA="ifAny")
ok.units.number <- t$UNITS.NUMBER == 1
sum(!ok.units.number)
```

## Constraint on Geocoding

Ranges
```{r}
summary(t$G.LATITUDE)
summary(t$G.LONGITUDE)
ok.geocoding <- t$G.LATITUDE != 0 & t$G.LONGITUDE != 0
sum(!ok.geocoding)
```

# Effect of all constraints

```{r}
totalExcluded <- 
  sum(!ok.sale.amount) +
  sum(!ok.document.type.code) +
  sum(!ok.transaction.type.code) +
  sum(!ok.sale.code) +
  sum(!ok.one.parcel) +
  sum(!ok.number.of.buildings) +
  sum(!ok.assessed.value) +
  sum(!ok.land.square.footage) +
  sum(!ok.universal.building.square.feet) +
  sum(!ok.living.square.feet) +
  sum(!ok.year.built) +
  sum(!ok.effective.year.built) +
  sum(!ok.total.rooms) +
  sum(!ok.units.number) +
  sum(!ok.geocoding)
totalExcluded
```
Function to write a text file
```{r}
table2Lines <- list()

Header2 <- function() {
  table2Lines <<- list("\\begin{verbatim}")
  table2Lines <<- c(table2Lines, "NUMBER OF OBSERVATIONS EXCLUDED BY EACH CRITERION IN ISOLATION")
  table2Lines <<- c(table2Lines, " ")
  line <- sprintf("%45s %12s", "Constraining Feature", "Num Excluded")
  table2Lines <<- c(table2Lines, line)
}

Data2 <- function(constraint, nExcluded) {
  line <- sprintf("%45s %12d", constraint, nExcluded)  
  invisible(table2Lines <<- c(table2Lines, line))
}

Footer2 <- function() {
  table2Lines <<- c(table2Lines, "\\end{verbatim}")
}

# These functions build the table
IfOnly <- function(constraintName, selectionVar) {
  excluded <- sum(!selectionVar)
  Data2(constraintName, excluded)
}

IfOnlyNum <- function(constraintName, num) {
  Data2(constraintName, num)
}
```

Generate the file.

```{r}
Header2()
IfOnly("SALE AMOUNT", ok.sale.amount)
IfOnly("DOCUMENT TYPE CODE", ok.document.type.code)
IfOnly("TRANSACTION TYPE CODE", ok.transaction.type.code)
IfOnly("SALE CODE", ok.sale.code)
IfOnly("one parcel", ok.one.parcel)
IfOnly("NUMBER OF BUILDINGS", ok.number.of.buildings)
IfOnly("assessed value", ok.assessed.value)
IfOnly("LAND SQUARE FOOTAGE", ok.land.square.footage)
IfOnly("UNIVERSAL BUILDING SQUARE FEET", ok.universal.building.square.feet)
IfOnly("LIVING SQUARE FEET", ok.living.square.feet)
IfOnly("YEAR BUILT", ok.year.built)
IfOnly("EFFECTIVE YEAR BUILT", ok.effective.year.built)
IfOnly("TOTAL ROOMS", ok.total.rooms)
IfOnly("UNITS NUMBER", ok.units.number)
IfOnly("geocoding", ok.geocoding)
IfOnlyNum("*TOTAL EXCLUDED*", totalExcluded)
Footer2()
```

Display file in HTML.
```{r}
str(table2Lines)
```

Write file.
```{r}
writeLines(as.character(table2Lines), con=path.excluded)
```

# Write the subset

```{r}
# check lengths for debugging
length(ok.sale.amount)
length(ok.document.type.code)
length(ok.transaction.type.code)
length(ok.sale.code)
length(ok.one.parcel)
length(ok.number.of.buildings)
length(ok.assessed.value)
length(ok.land.square.footage)
length(ok.universal.building.square.feet)
length(ok.living.square.feet)
length(ok.year.built)
length(ok.effective.year.built)
length(ok.total.rooms)
length(ok.geocoding)

# retain obs that are OK on all criteria
ok <- 
  ok.sale.amount &
  ok.document.type.code &
  ok.transaction.type.code &
  ok.sale.code &
  ok.one.parcel &
  ok.number.of.buildings &
  ok.assessed.value &
  ok.land.square.footage &
  ok.universal.building.square.feet &
  ok.living.square.feet &
  ok.year.built &
  ok.effective.year.built &
  ok.total.rooms &
  ok.geocoding
 
# give X a better name
t$transaction.obs.number <- t$X
t$X <- NULL

s <- t[ok,]
nodups <- unique(s)

nrow(t)       # number of transactions
nrow(s)       # number of observations in sample
nrow(nodups)  # number of non-duplicate observations in sample


# check that the de-duplication had no effect
if (nrow(s) != nrow(nodups))
  stop("de-duping found something to eliminate")

summary(nodups)

write.csv(nodups, path.output)
```
